---
title: "Firm Formalization and Electoral Participation -- Analyses and Replication Materials"
author: "Catarina Roman"
date: \today
output: pdf_document
---

```{r}
library(dplyr)
library(tidyr)
spallest <- read.csv("spallest.csv")
```

I will have to make 2 splits of the data (2 different panels): one for the intervals between federal/state elections and one for municipal elections.

The way to split up the analyses and make them tractable at first is splitting between states, and then unifying at the end. I will start with SP.


```{r}
# loading election result datasets

# municipal elections 
spel00 <- read.csv2('election-results/detalhe_votacao_munzona_2000_BRASIL.csv')
spel04 <- read.csv2('election-results/detalhe_votacao_munzona_2004_BRASIL.csv')
spel08 <- read.csv2('election-results/detalhe_votacao_munzona_2008_BRASIL.csv')
spel12 <- read.csv2('election-results/detalhe_votacao_munzona_2012_BRASIL.csv') 
spel16 <- read.csv2('election-results/detalhe_votacao_munzona_2016_BRASIL.csv') 
spel20 <- read.csv2('election-results/detalhe_votacao_munzona_2020_BRASIL.csv') 

# federal and state elections
spel02 <- read.csv2('election-results/detalhe_votacao_munzona_2002_BRASIL.csv')
spel06 <- read.csv2('election-results/detalhe_votacao_munzona_2006_BRASIL.csv')
spel10 <- read.csv2('election-results/detalhe_votacao_munzona_2010_BRASIL.csv')
spel14 <- read.csv2('election-results/detalhe_votacao_munzona_2014_BRASIL.csv')
spel18 <- read.csv2('election-results/detalhe_votacao_munzona_2018_BRASIL.csv')
#spel22




# Cleaning and splitting data between Executive and Legislative elections (split only necessary for null/blank votes -- turnout is the same across both since voters cast both ballots in one go)


## Municipal Elections

spel00 <- spel00 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel00p <- subset(spel00, spel00$DS_CARGO=="Prefeito")
spel00v <- subset(spel00, spel00$DS_CARGO=="Vereador") 

spel00p <- spel00p %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel00v <- spel00v %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))


spel04 <- spel04 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel04p <- subset(spel04, spel04$DS_CARGO=="Prefeito")
spel04v <- subset(spel04, spel04$DS_CARGO=="Vereador") 

spel04p <- spel04p %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel04v <- spel04v %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))


spel08 <- spel08 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel08p <- subset(spel08, spel08$DS_CARGO=="Prefeito")
spel08v <- subset(spel08, spel08$DS_CARGO=="Vereador") 

spel08p <- spel08p %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel08v <- spel08v %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))


#2012

spel12 <- spel12 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel12p <- subset(spel12, spel12$DS_CARGO=="Prefeito")
spel12v <- subset(spel12, spel12$DS_CARGO=="Vereador") 

spel12p <- spel12p %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel12v <- spel12v %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

#2016

spel16 <- spel16 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel16p <- subset(spel16, spel16$DS_CARGO=="Prefeito")
spel16v <- subset(spel16, spel16$DS_CARGO=="Vereador") 

spel16p <- spel16p %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel16v <- spel16v %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

#2020

spel20 <- spel20 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel20p <- subset(spel20, spel20$DS_CARGO=="Prefeito")
spel20v <- subset(spel20, spel20$DS_CARGO=="Vereador") 

spel20p <- spel20p %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel20v <- spel20v %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

## State Elections

spel02 <- spel02 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel02g <- subset(spel02, spel02$DS_CARGO=="Governador")
spel02de <- subset(spel02, spel02$DS_CARGO=="Deputado Estadual") 

spel02g <- spel02g %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel02de <- spel02de %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel06 <- spel06 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel06g <- subset(spel06, spel06$DS_CARGO=="Governador")
spel06de <- subset(spel06, spel06$DS_CARGO=="Deputado Estadual") 

spel06g <- spel06g %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel06de <- spel06de %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))


spel10 <- spel10 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel10g <- subset(spel10, spel10$DS_CARGO=="Governador")
spel10de <- subset(spel10, spel10$DS_CARGO=="Deputado Estadual") 

spel10g <- spel10g %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel10de <- spel10de %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

#2014

spel14 <- spel14 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel14g <- subset(spel14, spel14$DS_CARGO=="Governador")
spel14de <- subset(spel14, spel14$DS_CARGO=="Deputado Estadual") 

spel14g <- spel14g %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel14de <- spel14de %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))



#2018

spel18 <- spel18 %>% subset(NR_TURNO == 1) %>% select(CD_MUNICIPIO, DS_CARGO, QT_APTOS, QT_COMPARECIMENTO, QT_ABSTENCOES, QT_VOTOS_BRANCOS, QT_VOTOS_NULOS)

spel18g <- subset(spel18, spel18$DS_CARGO=="Governador")
spel18de <- subset(spel18, spel18$DS_CARGO=="Deputado Estadual") 

spel18g <- spel18g %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel18de <- spel18de %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))



## Federal Elections

spel02s <- subset(spel02, spel02$DS_CARGO=="Senador")
spel02df <- subset(spel02, spel02$DS_CARGO=="Deputado Federal") 

spel02s <- spel02s %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel02df <- spel02df %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))


spel06s <- subset(spel06, spel06$DS_CARGO=="Senador")
spel06df <- subset(spel06, spel06$DS_CARGO=="Deputado Federal") 

spel06s <- spel06s %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel06df <- spel06df %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel10s <- subset(spel10, spel10$DS_CARGO=="Senador")
spel10df <- subset(spel10, spel10$DS_CARGO=="Deputado Federal") 

spel10s <- spel10s %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel10df <- spel10df %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

#2014

spel14s <- subset(spel14, spel14$DS_CARGO=="Senador")
spel14df <- subset(spel14, spel14$DS_CARGO=="Deputado Federal") 

spel14s <- spel14s %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel14df <- spel14df %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

#2018

spel18s <- subset(spel18, spel18$DS_CARGO=="Senador")
spel18df <- subset(spel18, spel18$DS_CARGO=="Deputado Federal") 

spel18s <- spel18s %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))

spel18df <- spel18df %>% 
  select(!DS_CARGO) %>% 
  group_by(CD_MUNICIPIO) %>% 
  summarise(across(everything(), sum))


# V8 is codigo do municipio
#V12 is cargo
#V16 is QT_APTOS
#V18 is comparecimento 
#V19 is abstencoes 
#V20 is votos nominais
#V21 is brancos
#V22 is nulos
#V23 is legenda
```    


# BUILDING OUTCOME VARIABLES

```{r}

# Building OUTCOME variables


## 2000


# turnout and legislative non-votes
spel00v <- spel00v %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonleg = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel00p <- spel00p %>% mutate(nonexec = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


## 2004

# turnout and legislative non-votes
spel04v <- spel04v %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonleg = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel04p <- spel04p %>% mutate(nonexec = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


## 2008

# turnout and legislative non-votes
spel08v <- spel08v %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonleg = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel08p <- spel08p %>% mutate(nonexec = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

## 2012

# turnout and legislative non-votes
spel12v <- spel12v %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonleg = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel12p <- spel12p %>% mutate(nonexec = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

## 2016

# turnout and legislative non-votes
spel16v <- spel16v %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonleg = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel16p <- spel16p %>% mutate(nonexec = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

## 2020

# turnout and legislative non-votes
spel20v <- spel20v %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonleg = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel20p <- spel20p %>% mutate(nonexec = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


## 2002

# State Elections

# turnout and legislative non-votes
spel02de <- spel02de %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonde = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel02g <- spel02g %>% mutate(nong = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


# Federal Elections

#lower chamber non-votes
spel02df <- spel02df%>% mutate(nondf = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# upper chamber non-votes
spel02s <- spel02s %>% mutate(nons = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)



## 2006

# State Elections

# turnout and legislative non-votes
spel06de <- spel06de %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonde = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel06g <- spel06g %>% mutate(nong = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


# Federal Elections

#lower chamber non-votes
spel06df <- spel06df%>% mutate(nondf = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# upper chamber non-votes
spel06s <- spel06s %>% mutate(nons = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


## 2010

# State Elections

# turnout and legislative non-votes
spel10de <- spel10de %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonde = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel10g <- spel10g %>% mutate(nong = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


# Federal Elections

#lower chamber non-votes
spel10df <- spel10df%>% mutate(nondf = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# upper chamber non-votes
spel10s <- spel10s %>% mutate(nons = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

## 2014

# State Elections

# turnout and legislative non-votes
spel14de <- spel14de %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonde = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel14g <- spel14g %>% mutate(nong = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


# Federal Elections

#lower chamber non-votes
spel14df <- spel14df %>% mutate(nondf = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# upper chamber non-votes
spel14s <- spel14s %>% mutate(nons = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

## 2018

# State Elections

# turnout and legislative non-votes
spel18de <- spel18de %>% mutate(turnout = QT_COMPARECIMENTO/QT_APTOS) %>% mutate(nonde = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# executive non-votes
spel18g <- spel18g %>% mutate(nong = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)


# Federal Elections

#lower chamber non-votes
spel18df <- spel18df%>% mutate(nondf = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)

# upper chamber non-votes
spel18s <- spel18s %>% mutate(nons = (QT_VOTOS_BRANCOS + QT_VOTOS_NULOS)/QT_COMPARECIMENTO)
```




# Descriptive Stats

```{r}

# Calculating the EXPLANATORY VARIABLE: Difference between number of firms between two elections or stock of firms before an election


spallest <- spallest %>% 
  separate(V11, into=c('startyr', 'startmo', 'startday'), sep=c(4,6))
spallest$styrmo <- paste(spallest$startyr, spallest$startmo)
spallest$styrmo <- gsub(" ", "", spallest$styrmo)
spallest$startyr <- as.numeric(spallest$startyr)
spallest$startyr <- ifelse(spallest$startyr==0, NA, spallest$startyr)


spallest$newsimples <- ifelse(spallest$startyr==spallest$simpstyr, spallest$startyr, NA)
spallest$newsimplesind <- ifelse(spallest$startyr==spallest$simpstyr, 1,0)


spallest <- spallest %>% 
  separate(V7, into=c('cadyr', 'cadmo', 'cadday'), sep=c(4,6))
spallest$cadyrmo <- paste(spallest$cadyr, spallest$cadmo)
spallest$cadyrmo <- gsub(" ", "", spallest$styrmo)
spallest$sitcad <- ifelse(spallest$V6==2 | spallest$V6==3 | spallest$V6==4, 1, 0)

# THIS IS ALL FIRMS
# to do the same for only mei, all i have to do is subset the dataset by mei and re run the same code, just saving the individual datasets under different object names

spallest$meistyr <- as.numeric(spallest$meistyr)
spallest$meistyr <- ifelse(spallest$meistyr == 0, NA, spallest$meistyr)
spallest$mei <- ifelse(spallest$meistyr>0, "MEI", NA)  
spallest$MEI <- ifelse(spallest$mei=="MEI", 1, 0)
spallest$MEI <- ifelse(is.na(spallest$MEI), 0, spallest$MEI)
spallest$MEI <- as.factor(spallest$MEI)

```



```{r}

library(ggplot2)

meibyyear <- ggplot(spallest, aes(x=meistyr)) + geom_histogram()
meibyyear

# Every MEI is SIMPLES
# Not every SIMPLES is MEI
# So in 2007 there was that huge surge in SIMPLES subscriptions that were necessarily SIMPLES, not MEI because MEI did not exist
# But it is likely companies that already existed previously and just registered to SIMPLES because it decreased their bureaucracy / taxload

# I am only interested in firms that go from informal to formal, not in firms that were already formal and just adopted this tax scheme.


spallest$newsimples <- ifelse(spallest$startyr==spallest$simpstyr, spallest$startyr, NA)
spallest$newsimplesind <- ifelse(spallest$startyr==spallest$simpstyr, 1,0)

newfirmsperyear <- ggplot(spallest, aes(x=startyr, fill=MEI)) + geom_histogram() + xlim(2000, 2024) + stat_bin(bins=25)
newfirmsperyear

newfirmspost2008 <- ggplot(spallest, aes(x=startyr, fill=MEI)) + geom_histogram() + xlim(2007, 2024) 

newpersimples <- ggplot(spallest, aes(x=newsimples)) + geom_histogram() + xlim(2006, 2024)
newpersimples


# I am going to assume the MEI as a proxy for a firm that would otherwise be informal
# I need to figure out how to validate this

# Next step: remove from this MEI count the firms that are people being hired as freelance agents by larger firms so the latter avoid paying worker's rights (like the 13th salary) and additional taxes.
```









# BUILDING EXPLANATORY VARIABLES

```{r}

# FIX ALL OF THESE BY styrmo and cadyrmo

### MUNICIPAL ELECTIONS

spallest$flag_stock_00 <- ifelse(spallest$yrmo<200010, 1, 0)
spallest$flag2_stock_00 <- ifelse(spallest$cadyr<2001 & spallest$sitcad==0, -1, 0)
spallest$count_stock_00 <- spallest$flag2_stock_00+spallest$flag_stock_00
stock00 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_00 = sum(count_stock_00)) %>% 
  ungroup()

spallest$flag_delta_04 <- ifelse(spallest$startyr>2000 & spallest$startyr<2005, 1, 0)
spallest$flag2_delta_04 <- ifelse(spallest$startyr>2000 & spallest$startyr<2005 & spallest$sitcad==0, -1, 0)
spallest$count_delta_04 <- spallest$flag2_delta_04+spallest$flag_delta_04
delta04 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_04 = sum(count_delta_04)) %>% 
  ungroup()

spallest$flag_stock_04 <- ifelse(spallest$startyr<2005, 1, 0)
spallest$flag2_stock_04 <- ifelse(spallest$cadyr<2005 & spallest$sitcad==0, -1, 0)
spallest$count_stock_04 <- spallest$flag2_stock_04+spallest$flag_stock_04
stock04 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_04 = sum(count_stock_04)) %>% 
  ungroup()

spallest$flag_delta_08 <- ifelse(spallest$startyr>2004 & spallest$startyr<2009, 1, 0)
spallest$flag2_delta_08 <- ifelse(spallest$startyr>2004 & spallest$startyr<2009 & spallest$sitcad==0, -1, 0)
spallest$count_delta_08 <- spallest$flag2_delta_08+spallest$flag_delta_08
delta08 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_08 = sum(count_delta_08)) %>% 
  ungroup()

spallest$flag_stock_08 <- ifelse(spallest$startyr<2009, 1, 0)
spallest$flag2_stock_08 <- ifelse(spallest$cadyr<2009 & spallest$sitcad==0, -1, 0)
spallest$count_stock_08 <- spallest$flag2_stock_08+spallest$flag_stock_08
stock08 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_08 = sum(count_stock_08)) %>% 
  ungroup()

spallest$flag_delta_12 <- ifelse(spallest$startyr>2008 & spallest$startyr<2013, 1, 0)
spallest$flag2_delta_12 <- ifelse(spallest$startyr>2008 & spallest$startyr<2013 & spallest$sitcad==0, -1, 0)
spallest$count_delta_12 <- spallest$flag2_delta_12+spallest$flag_delta_12
spallest$mei_count_delta_12 <- ifelse(spallest$count_delta_12==1 & spallest$MEI==1, 1, 0)
delta12 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_12 = sum(count_delta_12), mei_delta_12 = sum(mei_count_delta_12)) %>% 
  ungroup()

spallest$flag_stock_12 <- ifelse(spallest$startyr<2013, 1, 0)
spallest$flag2_stock_12 <- ifelse(spallest$cadyr<2013 & spallest$sitcad==0, -1, 0)
spallest$count_stock_12 <- spallest$flag2_stock_12+spallest$flag_stock_12
spallest$mei_count_stock_12 <- ifelse(spallest$count_stock_12==1 & spallest$MEI==1, 1, 0)
stock12 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_12 = sum(count_stock_12), mei_stock_12 = sum(mei_count_stock_12)) %>% 
  ungroup()


spallest$flag_delta_16 <- ifelse(spallest$startyr>2012 & spallest$startyr<2017, 1, 0)
spallest$flag2_delta_16 <- ifelse(spallest$startyr>2012 & spallest$startyr<2017 & spallest$sitcad==0, -1, 0)
spallest$count_delta_16 <- spallest$flag2_delta_16+spallest$flag_delta_16
spallest$mei_count_delta_16 <- ifelse(spallest$count_delta_16==1 & spallest$MEI==1, 1, 0)
delta16 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_16 = sum(count_delta_16), mei_delta_16 = sum(mei_count_delta_16)) %>% 
  ungroup()

spallest$flag_stock_16 <- ifelse(spallest$startyr<2013, 1, 0)
spallest$flag2_stock_16 <- ifelse(spallest$cadyr<2013 & spallest$sitcad==0, -1, 0)
spallest$count_stock_16 <- spallest$flag2_stock_16+spallest$flag_stock_16
spallest$mei_count_stock_16 <- ifelse(spallest$count_stock_16==1 & spallest$MEI==1, 1, 0)
stock16 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_16 = sum(count_stock_16), mei_stock_16 = sum(mei_count_stock_16)) %>% 
  ungroup()

spallest$flag_delta_20 <- ifelse(spallest$startyr>2016 & spallest$startyr<2021, 1, 0)
spallest$flag2_delta_20 <- ifelse(spallest$startyr>2016 & spallest$startyr<2021 & spallest$sitcad==0, -1, 0)
spallest$count_delta_20 <- spallest$flag2_delta_20+spallest$flag_delta_20
spallest$mei_count_delta_20 <- ifelse(spallest$count_delta_20==1 & spallest$MEI==1, 1, 0)
delta20 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_20 = sum(count_delta_20), mei_delta_20 = sum(mei_count_delta_20)) %>% 
  ungroup()

spallest$flag_stock_20 <- ifelse(spallest$startyr<2021, 1, 0)
spallest$flag2_stock_20 <- ifelse(spallest$cadyr<2021 & spallest$sitcad==0, -1, 0)
spallest$count_stock_20 <- spallest$flag2_stock_20+spallest$flag_stock_20
spallest$mei_count_stock_20 <- ifelse(spallest$count_stock_20==1 & spallest$MEI==1, 1, 0)
stock20 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_20 = sum(count_stock_20), mei_stock_20 = sum(mei_count_stock_20)) %>% 
  ungroup()



### STATE/FEDERAL ELECTIONS

spallest$flag_stock_02 <- ifelse(spallest$startyr<2003, 1, 0)
spallest$flag2_stock_02 <- ifelse(spallest$cadyr<2003 & spallest$sitcad==0, -1, 0)
spallest$count_stock_02 <- spallest$flag2_stock_02+spallest$flag_stock_02
stock02 <- spallest %>% 
  group_by(V21) %>%
  summarise(stock_02 = sum(count_stock_02)) %>% 
  ungroup()

spallest$flag_delta_06 <- ifelse(spallest$startyr>2002 & spallest$startyr<2007, 1, 0)
spallest$flag2_delta_06 <- ifelse(spallest$startyr>2002 & spallest$startyr<2007 & spallest$sitcad==0, -1, 0)
spallest$count_delta_06 <- spallest$flag2_delta_06+spallest$flag_delta_06
delta06 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_06 = sum(count_delta_06)) %>% 
  ungroup()

spallest$flag_stock_06 <- ifelse(spallest$startyr<2007, 1, 0)
spallest$flag2_stock_06 <- ifelse(spallest$cadyr<2007 & spallest$sitcad==0, -1, 0)
spallest$count_stock_06 <- spallest$flag2_stock_06+spallest$flag_stock_06
stock06 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_06 = sum(count_stock_06)) %>% 
  ungroup()

spallest$flag_delta_10 <- ifelse(spallest$startyr>2006 & spallest$startyr<2011, 1, 0)
spallest$flag2_delta_10 <- ifelse(spallest$startyr>2006 & spallest$startyr<2011 & spallest$sitcad==0, -1, 0)
spallest$count_delta_10 <- spallest$flag2_delta_10+spallest$flag_delta_10
delta10 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_10 = sum(count_delta_10)) %>% 
  ungroup()

spallest$flag_stock_10 <- ifelse(spallest$startyr<2011, 1, 0)
spallest$flag2_stock_10 <- ifelse(spallest$cadyr<2011 & spallest$sitcad==0, -1, 0)
spallest$count_stock_10 <- spallest$flag2_stock_10+spallest$flag_stock_10
stock10 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_10 = sum(count_stock_10)) %>% 
  ungroup()

spallest$flag_delta_14 <- ifelse(spallest$startyr>2010 & spallest$startyr<2015, 1, 0)
spallest$flag2_delta_14 <- ifelse(spallest$startyr>2010 & spallest$startyr<2015 & spallest$sitcad==0, -1, 0)
spallest$count_delta_14 <- spallest$flag2_delta_14+spallest$flag_delta_14
spallest$mei_count_delta_14 <- ifelse(spallest$count_delta_14==1 & spallest$MEI==1, 1, 0)
delta14 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_14 = sum(count_delta_14), mei_delta_14 = sum(mei_count_delta_14)) %>% 
  ungroup()

spallest$flag_stock_14 <- ifelse(spallest$startyr<2015, 1, 0)
spallest$flag2_stock_14 <- ifelse(spallest$cadyr<2015 & spallest$sitcad==0, -1, 0)
spallest$count_stock_14 <- spallest$flag2_stock_14+spallest$flag_stock_14
spallest$mei_count_stock_14 <- ifelse(spallest$count_stock_14==1 & spallest$MEI==1, 1, 0)
stock14 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_14 = sum(count_stock_14), mei_stock_14 = sum(mei_count_stock_14)) %>% 
  ungroup()


spallest$flag_delta_18 <- ifelse(spallest$startyr>2014 & spallest$startyr<2019, 1, 0)
spallest$flag2_delta_18 <- ifelse(spallest$startyr>2014 & spallest$startyr<2019 & spallest$sitcad==0, -1, 0)
spallest$count_delta_18 <- spallest$flag2_delta_18+spallest$flag_delta_18
spallest$mei_count_delta_18 <- ifelse(spallest$count_delta_18==1 & spallest$MEI==1, 1, 0)
delta18 <- spallest %>% 
  group_by(V21) %>% 
  summarise(delta_18 = sum(count_delta_18), mei_delta_18 = sum(mei_count_delta_18)) %>% 
  ungroup()

spallest$flag_stock_18 <- ifelse(spallest$startyr<2019, 1, 0)
spallest$flag2_stock_18 <- ifelse(spallest$cadyr<2019 & spallest$sitcad==0, -1, 0)
spallest$count_stock_18 <- spallest$flag2_stock_18+spallest$flag_stock_18
spallest$mei_count_stock_18 <- ifelse(spallest$count_stock_18==1 & spallest$MEI==1, 1, 0)
stock18 <- spallest %>% 
  group_by(V21) %>% 
  summarise(stock_18 = sum(count_stock_18), mei_stock_18 = sum(mei_count_stock_18)) %>% 
  ungroup()
```


The explanatory variable I have in mind is what I call the RFEI, or the _Rate of Formalizing Entrepreneurial Initiative_. For now, the best quantification I can think of for this variable is $\frac{\Delta MEI_t}{Firm Stock_{t-1}}$.


```{r}
## BUILDING THE MUNICIPAL ELECTIONS PANEL

# 2000, 2004, 2008, 2012, 2016, 2020

# Legislative Elections (City Council)

spel00v$e2000 <- 1
spel00v <- spel00v %>% select(CD_MUNICIPIO, turnout, nonleg, e2000)
spel00v$e2004 <- 0
spel00v$e2008 <- 0
spel00v$e2012 <- 0
spel00v$e2016 <- 0
spel00v$e2020 <- 0

spel04v$e2004 <- 1
spel04v <- spel04v %>% select(CD_MUNICIPIO, turnout, nonleg, e2004)
spel04v$e2000 <- 0
spel04v$e2008 <- 0
spel04v$e2012 <- 0
spel04v$e2016 <- 0
spel04v$e2020 <- 0

spel08v$e2008 <- 1
spel08v <- spel08v %>% select(CD_MUNICIPIO, turnout, nonleg, e2008)
spel08v$e2000 <- 0
spel08v$e2004 <- 0
spel08v$e2012 <- 0
spel08v$e2016 <- 0
spel08v$e2020 <- 0

spel12v$e2012 <- 1
spel12v <- spel12v %>% select(CD_MUNICIPIO, turnout, nonleg, e2012)
spel12v$e2000 <- 0
spel12v$e2004 <- 0
spel12v$e2008 <- 0
spel12v$e2016 <- 0
spel12v$e2020 <- 0

spel16v$e2016 <- 1
spel16v <- spel16v %>% select(CD_MUNICIPIO, turnout, nonleg, e2016)
spel16v$e2000 <- 0
spel16v$e2004 <- 0
spel16v$e2008 <- 0
spel16v$e2012 <- 0
spel16v$e2020 <- 0

spel20v$e2020 <- 1
spel20v <- spel20v %>% select(CD_MUNICIPIO, turnout, nonleg, e2020)
spel20v$e2000 <- 0
spel20v$e2004 <- 0
spel20v$e2008 <- 0
spel20v$e2016 <- 0
spel20v$e2012 <- 0

munleg <- rbind(spel00v, spel04v, spel08v, spel12v, spel16v, spel20v)
muni <- read.csv2("municipios.csv")
muni <- muni %>% select(cdtom,cdibge)
muni2 <- read.csv('municipios_brasileiros_tse.csv')
muni2 <- muni2 %>% select(codigo_tse, codigo_ibge)
muni3 <- left_join(muni, muni2, by=c('cdibge'='codigo_ibge'))

munleg <- left_join(munleg, muni3, by=c('CD_MUNICIPIO'='codigo_tse'))

stock00$e2000 <- 1
stock04$e2004 <- 1
stock08$e2008 <- 1
stock12$e2012 <- 1
stock16$e2016 <- 1
stock20$e2020 <- 1

delta04$e2004 <- 1
delta08$e2008 <- 1 # between 2004 and 2008
delta12$e2012 <- 1
delta16$e2016 <- 1
delta20$e2020 <- 1


stock00$V21 <- as.integer(stock00$V21)
munleg1 <- left_join(munleg, stock00, by=c('cdtom'='V21', 'e2000'='e2000'))

stock04$V21 <- as.integer(stock04$V21)
munleg1 <- left_join(munleg1, stock04, by=c('cdtom'='V21', 'e2004'='e2004'))

stock08$V21 <- as.integer(stock08$V21)
munleg1 <- left_join(munleg1, stock08, by=c('cdtom'='V21', 'e2008'='e2008'))

stock12$V21 <- as.integer(stock12$V21)
munleg1 <- left_join(munleg1, stock12, by=c('cdtom'='V21', 'e2012'='e2012'))

stock16$V21 <- as.integer(stock16$V21)
munleg1 <- left_join(munleg1, stock16, by=c('cdtom'='V21', 'e2016'='e2016'))

stock20$V21 <- as.integer(stock20$V21)
munleg1 <- left_join(munleg1, stock20, by=c('cdtom'='V21', 'e2020'='e2020'))


delta04$V21 <- as.integer(delta04$V21)
munleg1 <- left_join(munleg1, delta04, by=c('cdtom'='V21', 'e2004'='e2004'))

delta08$V21 <- as.integer(delta08$V21)
munleg1 <- left_join(munleg1, delta08, by=c('cdtom'='V21', 'e2008'='e2008'))

delta12$V21 <- as.integer(delta12$V21)
munleg1 <- left_join(munleg1, delta12, by=c('cdtom'='V21', 'e2012'='e2012'))

delta16$V21 <- as.integer(delta16$V21)
munleg1 <- left_join(munleg1, delta16, by=c('cdtom'='V21', 'e2016'='e2016'))

delta20$V21 <- as.integer(delta20$V21)
munleg1 <- left_join(munleg1, delta20, by=c('cdtom'='V21', 'e2020'='e2020'))

munleg1[is.na(munleg1)] <- 0

munleg1$stock <- munleg1$stock_00 + munleg1$stock_04 + munleg1$stock_08 + munleg1$stock_12 + munleg1$stock_16 + munleg1$stock_20
munleg1$delta <- munleg1$delta_04 + munleg1$delta_08 + munleg1$stock_12 + munleg1$stock_16 + munleg1$stock_20
munleg1$mei_delta <- munleg1$mei_delta_12 +  munleg1$mei_delta_16 + munleg1$mei_delta_20
munleg1$mei_stock <- munleg1$mei_stock_12 + munleg1$mei_stock_16 + munleg1$mei_stock_20

#write.csv(munleg1, 'City council 2000-2020.csv')
```



```{r}

## BUILDING THE STATE ELECTIONS PANEL

# 2002, 2006, 2010, 2014, 2018

# Legislative Elections (State Legislature)

spel02de$e2002 <- 1
spel02de <- spel02de %>% select(CD_MUNICIPIO, turnout, nonde, e2002)
spel02de$e2006 <- 0
spel02de$e2010 <- 0
spel02de$e2014 <- 0
spel02de$e2018 <- 0

spel06de$e2006 <- 1
spel06de <- spel06de %>% select(V8, turnout, nonde, e2006)
spel06de$e2002 <- 0
spel06de$e2010 <- 0
spel06de$e2014 <- 0
spel06de$e2018 <- 0

spel06de <- spel06de %>% rename("CD_MUNICIPIO" = "V8")

spel10de$e2010 <- 1
spel10de <- spel10de %>% select(CD_MUNICIPIO, turnout, nonde, e2010)
spel10de$e2006 <- 0
spel10de$e2002 <- 0
spel10de$e2014 <- 0
spel10de$e2018 <- 0

spel14de$e2014 <- 1
spel14de <- spel14de %>% select(CD_MUNICIPIO, turnout, nonde, e2014)
spel14de$e2006 <- 0
spel14de$e2002 <- 0
spel14de$e2010 <- 0
spel14de$e2018 <- 0

spel18de$e2018 <- 1
spel18de <- spel18de %>% select(CD_MUNICIPIO, turnout, nonde, e2018)
spel18de$e2006 <- 0
spel18de$e2002 <- 0
spel18de$e2010 <- 0
spel18de$e2014 <- 0


stateleg <- rbind(spel02de, spel06de, spel10de, spel14de, spel18de)
stateleg <- left_join(stateleg, muni3, by=c('CD_MUNICIPIO'='codigo_tse'))

stock02$e2002 <- 1
stock06$e2006 <- 1
stock10$e2010 <- 1
stock14$e2014 <- 1
stock18$e2018 <- 1

delta06$e2006 <- 1
delta14$e2014 <- 1
delta18$e2018 <- 1
delta10$e2010 <- 1


stock02$V21 <- as.integer(stock02$V21)
stateleg1 <- left_join(stateleg, stock02, by=c('cdtom'='V21', 'e2002'='e2002'))

stock06$V21 <- as.integer(stock06$V21)
stateleg1 <- left_join(stateleg1, stock06, by=c('cdtom'='V21', 'e2006'='e2006'))

stock10$V21 <- as.integer(stock10$V21)
stateleg1 <- left_join(stateleg1, stock10, by=c('cdtom'='V21', 'e2010'='e2010'))

stock14$V21 <- as.integer(stock14$V21)
stateleg1 <- left_join(stateleg1, stock14, by=c('cdtom'='V21', 'e2014'='e2014'))

stock18$V21 <- as.integer(stock18$V21)
stateleg1 <- left_join(stateleg1, stock18, by=c('cdtom'='V21', 'e2018'='e2018'))



delta06$V21 <- as.integer(delta06$V21)
stateleg1 <- left_join(stateleg1, delta06, by=c('cdtom'='V21', 'e2006'='e2006'))

delta10$V21 <- as.integer(delta10$V21)
stateleg1 <- left_join(stateleg1, delta10, by=c('cdtom'='V21', 'e2010'='e2010'))

delta14$V21 <- as.integer(delta14$V21)
stateleg1 <- left_join(stateleg1, delta14, by=c('cdtom'='V21', 'e2014'='e2014'))

delta18$V21 <- as.integer(delta18$V21)
stateleg1 <- left_join(stateleg1, delta18, by=c('cdtom'='V21', 'e2018'='e2018'))

stateleg1[is.na(stateleg1)] <- 0

stateleg1$stock <- stateleg1$stock_02 + stateleg1$stock_06 + stateleg1$stock_10 + stateleg1$stock_14 + stateleg1$stock_18 
stateleg1$delta <- stateleg1$delta_06 + stateleg1$delta_10 + stateleg1$stock_14 + stateleg1$stock_18
stateleg1$mei_delta <- stateleg1$mei_delta_14 +  stateleg1$mei_delta_18 
stateleg1$mei_stock <- stateleg1$mei_stock_14 + stateleg1$mei_stock_18

# write.csv(stateleg1, 'State Legislature 2002-2018.csv')

```


```{r}

## BUILDING THE FEDERAL ELECTIONS PANEL

# 2002, 2006, 2010, 2014, 2018

# Legislative Elections (Lower Chamber National Congress)

spel02df$e2002 <- 1
spel02df <- spel02df %>% select(CD_MUNICIPIO, nondf, e2002)
spel02df$e2006 <- 0
spel02df$e2010 <- 0
spel02df$e2014 <- 0
spel02df$e2018 <- 0

spel06df$e2006 <- 1
spel06df <- spel06df %>% select(V8, nondf, e2006)
spel06df$e2002 <- 0
spel06df$e2010 <- 0
spel06df$e2014 <- 0
spel06df$e2018 <- 0

spel06df <- spel06df %>% rename("CD_MUNICIPIO" = "V8")

spel10df$e2010 <- 1
spel10df <- spel10df %>% select(CD_MUNICIPIO, nondf, e2010)
spel10df$e2006 <- 0
spel10df$e2002 <- 0
spel10df$e2014 <- 0
spel10df$e2018 <- 0

spel14df$e2014 <- 1
spel14df <- spel14df %>% select(CD_MUNICIPIO, nondf, e2014)
spel14df$e2006 <- 0
spel14df$e2002 <- 0
spel14df$e2010 <- 0
spel14df$e2018 <- 0

spel18df$e2018 <- 1
spel18df <- spel18df %>% select(CD_MUNICIPIO, nondf, e2018)
spel18df$e2006 <- 0
spel18df$e2002 <- 0
spel18df$e2010 <- 0
spel18df$e2014 <- 0


lower <- rbind(spel02df, spel06df, spel10df, spel14df, spel18df)
lower <- left_join(lower, muni3, by=c('CD_MUNICIPIO'='codigo_tse'))

stock02$e2002 <- 1
stock06$e2006 <- 1
stock10$e2010 <- 1
stock14$e2014 <- 1
stock18$e2018 <- 1

delta06$e2006 <- 1
delta14$e2014 <- 1
delta18$e2018 <- 1
delta10$e2010 <- 1


stock02$V21 <- as.integer(stock02$V21)
lower1 <- left_join(lower, stock02, by=c('cdtom'='V21', 'e2002'='e2002'))

stock06$V21 <- as.integer(stock06$V21)
lower1 <- left_join(lower1, stock06, by=c('cdtom'='V21', 'e2006'='e2006'))

stock10$V21 <- as.integer(stock10$V21)
lower1 <- left_join(lower1, stock10, by=c('cdtom'='V21', 'e2010'='e2010'))

stock14$V21 <- as.integer(stock14$V21)
lower1 <- left_join(lower1, stock14, by=c('cdtom'='V21', 'e2014'='e2014'))

stock18$V21 <- as.integer(stock18$V21)
lower1 <- left_join(lower1, stock18, by=c('cdtom'='V21', 'e2018'='e2018'))



delta06$V21 <- as.integer(delta06$V21)
lower1 <- left_join(lower1, delta06, by=c('cdtom'='V21', 'e2006'='e2006'))

delta10$V21 <- as.integer(delta10$V21)
lower1 <- left_join(lower1, delta10, by=c('cdtom'='V21', 'e2010'='e2010'))

delta14$V21 <- as.integer(delta14$V21)
lower1 <- left_join(lower1, delta14, by=c('cdtom'='V21', 'e2014'='e2014'))

delta18$V21 <- as.integer(delta18$V21)
lower1 <- left_join(lower1, delta18, by=c('cdtom'='V21', 'e2018'='e2018'))

lower1[is.na(lower1)] <- 0

lower1$stock <- lower1$stock_02 + lower1$stock_06 + lower1$stock_10 + lower1$stock_14 + lower1$stock_18 
lower1$delta <- lower1$delta_06 + lower1$delta_10 + lower1$stock_14 + lower1$stock_18
lower1$mei_delta <- lower1$mei_delta_14 +  lower1$mei_delta_18 
lower1$mei_stock <- lower1$mei_stock_14 + lower1$mei_stock_18

#write.csv(lower1, 'Congress Lower Chamber 2002-2018.csv')

```



